Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2021-01-25 xrb2/paper.tex
update conclusions

We ran a number of simulations of laterally propagating flames in XRBs in order to explore the effects of rotation and thermal structure. We found that increasing the rotation rate increased the energy generation rate within the flame and enhanced nuclear burning. Apart from the lowest rotation run (which failed to ignite), flame propagation was not noticeably impacted by rotation rate; by the time the different flame fronts reached steady propagation, they shared comparable velocities. These results are likely due to the rotation-dependent strength of the Coriolis force its confinement of the flame balancing the enhanced nuclear burning. We explored several models with different crustal temperatures to determine what effects mechanisms such as higher accretion rates, crustal heating and shallow heating may have on flame propagation. We found that increasing the temperature of the crust significantly enhanced the flame propagation. This we believe to be because a cooler crust allows heat to more efficiently be transferred away from the flame itself, therefore reducing the flame's temperature, slowing burning and consequently reducing its propagation speed. At higher temperatures, we saw that the flame itself accelerated, the increased temperature and inability for heat to be transported away from the flame front driving unstable, accelerating flame propagation. We saw that if the crust temperature was too high, then instead of a flame the entire atmosphere would burn steadily. This is reminiscent of what is seen for neutron stars with accretion rates exceeding the Eddington luminosity.

In future simulations, we would like to improve and expand our
simulations in order to better understand the processes at play and to
include more physics. This includes adding tracer particles to the
simulations so we can monitor the fluid motion and perform more
detailed nucleosynthesis; extending our simulations to 3D, which would
hopefully alleviate some of the boundary effects we have observed in
these simulations but will require significant computational resources;
and exploring the resolution of our simulations more so that we can
ensure that we have resolved all of the necessary physical processes. We
would also like to model H/He flames, as these are the sites of
rp-process nucleosynthesis~\citep{rpprocess}.  Initially we will use the same reaction
sequence explored in our previous convection
calculations~\citep{xrb2}.  Finally, we
recently added an MHD solver to \castro~\citep{sazo2020thesis}; this
will allow us in the future to explore the effects of magnetic fields on flame
propagation in XRBs.

>>>>

We ran a number of simulations of laterally propagating flames in XRBs in order to explore the effects of rotation and thermal structure. We found that increasing the rotation rate increased the energy generation rate within the flame and enhanced nuclear burning. Apart from the lowest rotation run (which failed to ignite), flame propagation was not noticeably impacted by rotation rate; by the time the different flame fronts reached steady propagation, they shared comparable velocities. These results are likely due to the rotation-dependent strength of the Coriolis force and its confinement of the flame balancing the enhanced nuclear burning. We explored several models with different crustal temperatures to determine what effects mechanisms such as higher accretion rates, crustal heating and shallow heating may have on flame propagation. We found that increasing the temperature of the crust significantly enhanced the flame propagation. This we believe to be because a cooler crust allows heat to more efficiently be transferred away from the flame itself, therefore reducing the flame's temperature, slowing burning and consequently reducing its propagation speed. At higher crustal temperatures, we saw that the inability for heat to be efficiently transported away from the flame front increased the flame temperature, driving unstable, accelerating flame propagation. We saw that if the crust temperature was too high, then instead of a flame the entire atmosphere would burn steadily. This is reminiscent of what is seen for neutron stars with accretion rates exceeding the Eddington luminosity.

In future work, we would like to improve and expand our
simulations in order to better understand the processes at play and to
include more physics. This includes adding tracer particles to the
simulations so we can monitor the fluid motion and perform more
detailed nucleosynthesis; extending our simulations to 3D, which would
hopefully alleviate some of the boundary effects we have observed in
these simulations but will require significant computational resources;
and exploring the resolution of our simulations more so that we can
ensure that we have resolved all of the necessary physical processes. We
would also like to model H/He flames, as these are the sites of
rp-process nucleosynthesis~\citep{rpprocess}.  Initially we will use the same reaction
sequence explored in our previous convection
calculations~\citep{xrb2}.  Finally, we
recently added an MHD solver to \castro~\citep{sazo2020thesis}; this
will allow us in the future to explore the effects of magnetic fields on flame
propagation in XRBs.
